---
title: " Daily gold pricesForecasting Models"
author: "Arinze Francis"
date: '2022-07-08'
output: html_document
---

# R set up
```{r}
rm(list=ls())
options(scipen=999,digits=4)
rm

```


# Load R packages

```{r}
library(readxl)
library(tidyquant)
library(tidyverse)
library(lubridate)
library(xts)
library(quantmod)
library(tseries)
library(zoo)
library(ggplot2)
library(fpp2)
library(data.table)
library(forecast)
library(anytime)
```


# Dataset Loading

```{r}
data <- ts(gold, start = c(1985,1), frequency = 365.25)
class(data)
frequency(data)
summary(data)

plot(data)
# There is an outlier, the long spike.
# There are NA's values, the broken plot which could be seen better by using summary NA functions.


```


# Cleaning Na's and outliers: Since I didn't get the missing values from the data collector; I will have to use a ts function to statistically get it.

```{r}
# Cleaning the data
data_clean <- tsclean(data)
data_clean

# Further check if there are NAs
summary(data_clean) 

# plot the cleaned data

plot.ts(data_clean,ylab="Morning Gold Prices", xlab="Daily", main="Morning Gold Prices vs   Individual days")
# There is heteroscedasticity
# mean is constant
# variance is not constant.
# There is no seasonality component


```
# To ascertain if there is a seasonal components

```{r}

ggseasonplot(data_clean, year.labels = T, year.labels.left = T) +ylab("degree") +ggtitle("Seasonal Plot:  Daily Gold Prices Data ")
# There is no seasonality


ggseasonplot(data_clean, polar = T) +ylab("degree") +ggtitle("Polar plot: Daily Gold Prices")
# The plot is circular meanaing there is no seasonality.


```

# Splitting the dataset for training and testing

```{r}

data_train <- window(data_clean, start=c(1985,1),end=c(1987,5), freq=365.25)
data_train


data_test <- window(data_clean, start=c(1987,6), freq=365.25)
data_test


autoplot(data_train, series = "Train") + autolayer(data_test, series = "Test") + ggtitle("Gold Prices train and test set") +xlab("Individual days") +ylab("Gold Prices")+guides(colour=guide_legend(title = "Forecast"))


```

# Data forecasting methods 


```{r}

naiveF <- naive(data_train, h=365*3) ; naiveF  # Naive method
plot(naiveF)


meanFo <- meanf(data_train, h=365*3);meanFo   # Average method
plot(meanFo)

rwff <- rwf(data_train, h=365*3, drift = T) ;rwff  # Random walk with drift







```
# Accuracy Measures of the forecasting Methods and checking residuals

```{r}

# Accuracy measures

accuracy(naiveF, data_test)
accuracy(meanFo, data_test)
accuracy(rwff, data_test)

# the Random walk with drift method has the best model becuase it has the lowest RMSE, MAE, MASE, MAPE


# Checking residuals

print(summary(meanFo))
checkresiduals(meanFo)

print(summary(naiveF))
checkresiduals(naiveF)

print(summary(rwff))
checkresiduals(rwff)


# the Random walk with drit method  has the lowest std.dev

```

# Further plotting


```{r}
naiveFF <- naive(data_clean, h=365*3) ; naiveFF  # Naive method
plot(naiveFF)

meanFoo <- meanf(data_clean, h=365*3);meanFoo  # Average method
plot(meanFoo)

rwfff <- rwf(data_clean, h=365*3, drift = T) ;rwfff  # Random walk with drift
plot(rwfff)


print(summary(meanFoo))
checkresiduals(meanFoo)

print(summary(naiveFF))
checkresiduals(naiveFF)

print(summary(rwfff))
checkresiduals(rwfff)



```



# Comaprison plot for 3 models

```{r}

library(ggplot2)

autoplot(data_clean) + forecast::autolayer(meanFoo$mean, series = "Average model")+forecast::autolayer(rwfff$mean, series = "Random Walk with drift model")+forecast::autolayer(naiveFF$mean, series = "Naive model")+xlab("Individual days") +ylab("Daily Gold Prices") + guides (colour = guide_legend(title = "Forecast Model Methods")) + theme(legend.position = c(0.9,0.9))

# Random walk with drift methods is the best model cos it picked up the trend

```









































